Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-18 bbbq_article.tex
Removed obsolete text
Add results separated by number of spanners

% \paragraph{Immune response by MHC-I}

% GvdB: Not really relevant. You don’t really need to explain the role of DCs. What is far more important is the concept of what an MHC is: a protein complex that contains a fold, called the peptide-binding groove, which presents peptides. Also immediately explain that MHC-I presents peptides of about 9-amino acids in length to cytolytic T cells, whereas MHC-II presents longer peptides to helper T cells.

% Antigen-presenting cells such as dendritic cells
% use the MHC class I molecules to 
% present randomly sampled peptides fragments,
% as created by the immunoproteasome (\cite{chapiro2006destructive}).
% If a cell gets infected by an intracellular pathogen, also the virus' peptides
% will also be presented at the cell surface. The foreign viral antigen is detected 
% by cytotoxic T lymphocytes, which will kill the infected cell.

% \paragraph{Immune response by MHC-II, contrast with APC}

% GvdB: Also not really relevant

% All pathogens can be detected by the foreign proteins on 
% their (bacterial or fungal) cell walls or (viral) envelope.
% For bacteria and fungi, this is the main mode of their detection,
% as these do not infect a cell with their (foreign) DNA.
% T and B cells express MHC-II to detect foreign peptide fragments.
% An immune response is started when MHC-II detects a pathogen.

@@

% \paragraph{MHC-I presents TMH-derived epitopes in humans more often}

%For MHC-I, it was found that predicted epitopes derived 
%from human transmembrane helices (TMHs)
%are over-presented by all 5 HLA-A and 
%most of 8 HLA-B super-types (\cite{bianchi2017}),
%due the high hydrophobicity as is required for a TMH to span
%the (hydrophobic) cell membrane, combined with the fact that the MHC
%binding cleft is hydrophobic as well.
%
%It is unknown if the presentation of hydrophobic/TMH residues
%is evolutionary selected for.
%One reason to expect a reduced variability (and hence 
%evolutionary conservation) in TMHs, is
%that these are restricted in their evolution
%by the functional requirement to span a lipid bilayer.
%Due to this, the TMHs that pathogens possess, 
%have a lower chance to develop an escape mutation,
%as many mutations will result in a dysfunctional TMH.

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_n_tmhs.png}
  \caption{
    Percentage of SNPs found in TMHs,
    separated per number of TMHs.
    The dashed diagonal lines show the expected linear trend line
    if TMHs and soluble protein regions are equally conserved.
    \richel{GvdB: As discussed, I would simply plot 1 and multiple (>1) TMHs.}
  }
  \label{fig:f_snps_found_and_expected_per_n_tmhs}
\end{figure}
)

We decided to split this analysis based on the number of TMHs
a protein has \richel{Why? Provide brief rationale again. Thus TMH-TMH interactions or something}. We hypothesize that single-spanners (i.e. proteins
with one TMH predicted) are less conserved, when compared to multi-spanners,
as single spanners just need to span a membrane, while multi-spanners
might have interactions between their TMHs, 
for example to accommodate active sites, and 
thus might have additional structural constraints.
As can be seen in figure \ref{fig:f_snps_found_and_expected_per_n_tmhs}, 
there is no clear pattern to be observed. 
\richel{GvdB: Rewrite after pooling >1 TMHs}

>>>>

% GvdB: Not really relevant. You don’t really need to explain the role of DCs. 
% What is far more important is the concept of what an MHC is: 
% a protein complex that contains a fold, called the peptide-binding groove, 
% which presents peptides. Also immediately explain that MHC-I 
% presents peptides of about 9-amino acids in length to cytolytic T cells, 
% whereas MHC-II presents longer peptides to helper T cells.

@@

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
  \caption{
    Percentage of SNPs found in TMHs per spanner.
    The dashed diagonal lines show the expected linear trend line
    if TMHs and soluble protein regions are equally conserved.
  }
  \label{fig:f_snps_found_and_expected_per_spanner}
\end{figure}
)

We decided to split this analysis based on the number of TMHs
a protein has \richel{Why? Provide brief rationale again. Thus TMH-TMH interactions or something}. We hypothesize that single-spanners (i.e. proteins
with one TMH predicted) are less conserved, when compared to multi-spanners,
as single spanners just need to span a membrane, while multi-spanners
might have interactions between their TMHs, 
for example to accommodate active sites, and 
thus might have additional structural constraints.
As can be seen in figure \ref{fig:f_snps_found_and_expected_per_spanner}, 
both single- and multi-spanners are evolutionarily conserved.
